NTM CT Scoring Agreement

### 2025-07-31 Stacey Martiniano/ntm_agreement.qmd
need to ask Stacey about the total score thing

## Total disease feature score (orange)

```{r}
# Import disease state
outcomes <-
  read.csv("./Stacey Martiniano/NTM CT Scoring/Data_Clean/disease.key.csv",
    na.strings = ""
  )
outcomes$NTM.PD..Y.N. <- factor(outcomes$NTM.PD..Y.N.,
  levels = c("No", "Yes"), ordered = T
)
outcomes <- outcomes %>%
  select(cff_registry_id, NTM.PD..Y.N.) %>%
  rename(`NTM-PD` = NTM.PD..Y.N.)
# For the duplicated scans, just use the first reading and combine with the
# singletons
scores <- rbind(data, data_dups[data_dups$scorer_repeat == 1, ])
# Add gold standards
scores <- rbind(scores, data_gs[data_gs$scorer == "GS", ])
# Calculate the scores for each feature.
# When calculating the scores, take the majority vote by scan, feature, and
# location, since that's how they scored it.
# For example: If one scorer detected atelectasis in the RUL, one in the RML,
# and one in the RLL, but the majority at each location was "absent," then that
# scan would have a 0 for atelectasis. In other words, score the locations by
# majority first, then add the locations together for the feature score.
feature_scores <- scores %>%
  group_by(scan_id, Feature, Location) %>%
  summarise(Majority = factor(sum(value == "Present"),
    levels = 0:3, labels = c("Absent", "Absent", "Present", "Present")
  ), .groups = "drop_last") %>%
  summarise(Score = sum(Majority == "Present"), .groups = "drop")
# Add outcome data
feature_scores <- left_join(feature_scores, outcomes,
  by = join_by(scan_id == cff_registry_id)
)
# Summary table
feature_scores %>%
  pivot_wider(names_from = Feature, values_from = Score) %>%
  select(-scan_id) %>%
  tbl_summary(by = `NTM-PD`, missing_text = "Missing")
feature_scores %>%
  pivot_wider(names_from = Feature, values_from = Score) %>%
  select(-scan_id) %>%
  tbl_summary(
    by = `NTM-PD`, missing_text = "Missing",
    type = list(
      c(
        "Atelectasis", "Bronchiectasis", , "Cavity", "Consolidation", "Cyst",
        "Ground Glass", "Mucoid Impaction", "Nodule", "Tree In Bud",
        "Tree In Bud OR Nodule"
      ) ~ "continuous2"
    ),
    statistic = all_continuous() ~ c("{median} ({p25}, {p75})", "{min}, {max}")
  )
```

>>>>

```{r}
# Import disease state
outcomes <-
  read.csv("./Stacey Martiniano/NTM CT Scoring/Data_Clean/disease.key.csv",
    na.strings = ""
  )
outcomes$NTM.PD..Y.N. <- factor(outcomes$NTM.PD..Y.N.,
  levels = c("No", "Yes"), ordered = T
)
outcomes <- outcomes %>%
  select(cff_registry_id, NTM.PD..Y.N.) %>%
  rename(`NTM-PD` = NTM.PD..Y.N.)
# For the duplicated scans, just use the first reading and combine with the
# singletons
scores <- rbind(data, data_dups[data_dups$scorer_repeat == 1, ])
# Add gold standards
scores <- rbind(scores, data_gs[data_gs$scorer == "GS", ])
```

## Feature score (orange)

```{r}
# Calculate the scores for each feature.
# When calculating the scores, take the majority vote by scan, feature, and
# location, since that's how they scored it.
# For example: If one scorer detected atelectasis in the RUL, one in the RML,
# and one in the RLL, but the majority at each location was "absent," then that
# scan would have a 0 for atelectasis. In other words, score the locations by
# majority first, then add the locations together for the feature score.
feature_scores <- scores %>%
  group_by(scan_id, Feature, Location) %>%
  summarise(Majority = factor(sum(value == "Present"),
    levels = 0:3, labels = c("Absent", "Absent", "Present", "Present")
  ), .groups = "drop_last") %>%
  summarise(Score = sum(Majority == "Present"), .groups = "drop")
# Add outcome data
feature_scores <- left_join(feature_scores, outcomes,
  by = join_by(scan_id == cff_registry_id)
)
# Summary table
feature_scores %>%
  pivot_wider(names_from = Feature, values_from = Score) %>%
  select(-scan_id) %>%
  tbl_summary(by = `NTM-PD`, missing_text = "Missing")
feature_scores %>%
  pivot_wider(names_from = Feature, values_from = Score) %>%
  select(-scan_id) %>%
  tbl_summary(
    by = `NTM-PD`, missing_text = "Missing",
    type = list(
      c(
        "Atelectasis", "Bronchiectasis", , "Cavity", "Consolidation", "Cyst",
        "Ground Glass", "Mucoid Impaction", "Nodule", "Tree In Bud",
        "Tree In Bud OR Nodule"
      ) ~ "continuous2"
    ),
    statistic = all_continuous() ~ c("{median} ({p25}, {p75})", "{min}, {max}")
  )
```

## Lobe score (pink)

```{r}
# Lobe scores similar to above but we'll add over the lobes rather than the
# features
lobe_scores <- scores %>%
  group_by(scan_id, Location, Feature) %>%
  summarise(Majority = factor(sum(value == "Present"),
    levels = 0:3, labels = c("Absent", "Absent", "Present", "Present")
  ), .groups = "drop_last") %>%
  summarise(Score = sum(Majority == "Present"), .groups = "drop")
# Add outcome data
lobe_scores <- left_join(lobe_scores, outcomes,
  by = join_by(scan_id == cff_registry_id)
)
# Summary table
lobe_scores %>%
  pivot_wider(names_from = Location, values_from = Score) %>%
  select(-scan_id) %>%
  tbl_summary(by = `NTM-PD`, missing_text = "Missing")
lobe_scores %>%
  pivot_wider(names_from = Location, values_from = Score) %>%
  select(-scan_id) %>%
  tbl_summary(
    by = `NTM-PD`, missing_text = "Missing",
    type = list(
      c("RUL", "RML", "RLL", "LUL", "Ling", "LLL") ~ "continuous2"
    ),
    statistic = all_continuous() ~ c("{median} ({p25}, {p75})", "{min}, {max}")
  )
```

## Total score (pink)

The way we are currently scoring CT scans, the lobe-wise and feature-wise sums are exactly the same.

```{r}
# Basically we are adding the two previous dataframes together
# lobe_scores = lobe_scores %>% filter()
# feature_scores = feature_scores %>% group_by(scan_id) %>%
#   summarise(Total = sum(Score))
```
